Model to study co-evolution of trade an culture

### 2015-04-06 article.tex
minor modifications to intro and ccl

%XRC this process played a role in the trajectories of cultural change of the different regions, which were interconnected by trade
The trade possibilities between provinces modify the transmission of culture within each province and between them. The influence can comes from two main sources. First, when a new connection is made between two provinces, new products can arrive in their respective markets. The arrival of new goods naturally changes the rate of adoption of every goods of the market CITE. Moreover, the arrival of new products can lead to their recombination with previous products and therefore the creation of an new item. This further modify the structure of the market of the province. Second, persons are also travelling on the roads of the trade network, each carrying its own culture. Large movements of populations may lead to changes in the cultures. For example, when Roman soldiers were stationing in a province, Roman products such as olive oil were delivered to that province CITE ROMAN DELIVERING OIL. This can lead to the adoption of olive oil in regions that were originally not using it.

%XRC these two elements (cultural transmission and trade networks) are traditionally been studied as isolated systems.
Traditionally cultural transmission and trade are studied distinctively.  On the one hand, the cultural transmission mechanisms studied do not take economical consideration into accounts. The works done are interested in the adoption of various cultural trait without taking into consideration their economical value. On the other hand, trading behaviours are studied within their own framework which are designed to measure the amount of goods exchanged and the value at which they are exchanged. Within these frameworks the cultural transmission between persons is not present.

%XRC this is wrong because blablabla
As it has been stated before, trades influence cultural transmission by modifying both the goods present in a region and the people consuming these goods. Certainly, analysing separately trade and cultural transmission allows to formalize their inner mechanisms and understand their respective dynamics. However, in order to analyse the adoption of goods within a province one must take into account both trading strategy and cultural transmission. For example, the modification of good's consumption after the rupture of a connection in the trade system is better explained by knowing the local adoption of this good (cultural transmission) and the prices of this good (trade).

%XRC we propose a model of coevolution to explore the deep links between trade and cultural transmission in the ancient Mediterranean
Ultimately, we aim to create a model to study the relation between cultural change and trade networks within the Roman Empire. Examples of questions that could be asked to our model are: What is the relation between distances in the trade network and cultural changes ? Did anti-conformism promote the adaptation of new products coming from remote provinces ? In order to answer these questions the model used has to take into account both cultural effects and trade influence. More precisely, the variations in these two aspects should be modelled and should influence each other.  Therefore, in order to answer the questions targeted we need a co-evolutionary model able to take into account both trade and cultural changes.

@@

The model presented in this work aims to offer a framework for the co-evolution of culture and trades which relies on a low number of parameters. Moreover, to be useful, the framework is developed with the aim to ease the comparison of results obtained when taking cultural or trade into consideration.  This article will focus on the presentation of the framework, its justification and more importantly its validation by reproducing previous results obtained on cultural change, and the ability to produce trade dynamics. The results obtained when studying trade and cultural change are compared with each other.

@@

This article has propose a framework to study both cultural transmission mechanisms and trade mechanisms. Our framework is based on a central element, the value vector, which is interpreted differently depending on the mechanisms studied. We have shown the validity of this approach by reproducing expected results on both the cultural and trade side. On the cultural transmission side we have shown that the implementation of the ``neutral model" leads to the expected observations on the variants of the vector value: a power law. When implementing trading mechanisms we observe the convergence of prices to the expected values and the improvement of the scores of the agents.

>>>>

%XRC this process played a role in the trajectories of cultural change of the different regions, which were interconnected by trade
Trade roads between provinces generates cultural changes due to two main effects. First, when a new connection is made between two provinces, new goods can arrive in their respective markets. The arrival of new goods naturally changes the rate of adoption of all the goods of the market CITE. Moreover, a good recently introduced in a province can be recombined with previous goods in order to create yet a new type of good, which further modify the structure of the market of the province. The second element affecting the change of culture within a province is due to the persons travelling on the roads of the trade network, each carrying its own culture. Large movements of populations may lead to changes in the cultures. For example, when Roman soldiers were stationing in a province, Roman products such as olive oil were delivered to that province CITE ROMAN DELIVERING OIL, which may have lead to the adoption of olive oil in regions that were originally not using it.

%XRC these two elements (cultural transmission and trade networks) are traditionally been studied as isolated systems.
Usually, the studies on cultural change mechanisms focus on the adoption of cultural goods such as pottery by a population. However, these studies do not take into account economical factors, such as the cost of production or acquisition of the goods. Moreover, the effect that trade may have on the frequency of goods in the population or in neighbourhood populations is not analysed. A dedicated research field studies the effect of trade mechanisms on the change of prices and the modification of goods repartition in the population. This field relies on a specific set of framework which do not take into account the results obtained on the study of cultural change mechanisms. Certainly, this division in the analysis of cultural change and trade and cultural transmission allows for a precise formalization of their inner mechanisms and therefore a good understanding of their respective dynamics.
%XRC this is wrong because blablabla
However, as it has been shown before, trades influence cultural change by modifying both the goods present in a region and the people consuming these goods.  Therefore, in order to analyse the adoption of goods within a province one must take into account both trading strategy and cultural transmission. For example, the modification of good's consumption after the rupture of a connection in the trade system is better explained by knowing the local adoption of this good (cultural change) and the prices of this good (trade).

%XRC we propose a model of coevolution to explore the deep links between trade and cultural transmission in the ancient Mediterranean
Ultimately, we aim to create a model to study the relation between cultural change and trade networks within the Roman Empire. Examples of questions that could be asked to our model are: What is the relation between distances in the trade network and cultural changes ? Did anti-conformism promote the adaptation of new products coming from remote provinces ? In order to answer these questions the model developed has to take into account both cultural effects and trade influence. More precisely, the variations in these two aspects should be modelled and should influence each other.  Therefore, in order to answer the questions targeted we need a co-evolutionary model able to take into account both trade and cultural changes.

@@

Therefore, in this work, we aim for the development of a framework to study the co-evolution of culture and trades which relies on a low number of parameters. The idea behind the development of such framework is to ease the comparison of different cultural change and trade mechanisms, which means that the framework should be flexible enough to allow for the implementation of both type of mechanisms and with enough structural elements to allow the comparison of various implementations. This article will focus on the presentation of the framework, its justification and more importantly its validation by reproducing previous results obtained on cultural change, and the ability to produce trade dynamics. The results obtained when studying trade and cultural change are compared with each other.

@@

This article has proposed a framework to study both cultural change mechanisms and trade mechanisms. The development of our framework was first aimed at simplicity which is achieved by the use of two vectors (quantity and value) and four processes (production, trade, cultural transmission and innovation). The second aim of the work conducted was to obtain a flexible framework which is possible since each of the processes can be implemented accordingly to the question studied. We have shown the validity of this approach by reproducing expected results on both the cultural and trade side. On the cultural transmission side we have shown that the implementation of a ``neutral model" leads to the expected observations on the variants of the vector value: a power law. When implementing trading mechanisms we observe the convergence of prices to the expected values and the improvement of the scores of the agents.
